\section{Fundamentals}

\subsection{Securities Trading and Markets}

The basic function of a market is to bring buyers and sellers together.

\begin{process}
\hlt{(Four Components of a Trade)}
\begin{enumerate}[label=\roman*.]
\setlength{\itemsep}{0pt}
\item Acquisition of information and quotes
\begin{enumerate}[label=\arabic*.]
\setlength{\itemsep}{0pt}
\item Quality information and transparency are crucial to price discovery
\item Transparent markets quickly disseminate high-quality information
\item Opaque markets are those that lack transparency
\end{enumerate}
\item Routing of the trade order
\begin{enumerate}[label=\arabic*.]
\setlength{\itemsep}{0pt}
\item Selecting the brokers to handle the trades
\item Deciding which markets will execute the trades and transmitting the trades to the markets
\end{enumerate}
\item Execution. Buys are matched and executed against sells according to the rules of the market
\item Confirmation, clearance and settlement
\begin{enumerate}[label=\arabic*.]
\setlength{\itemsep}{0pt}
\item Clearance is the recording and comparison of the trade records
\item Settlement involves the actual delivery of the security and its payment
\item Might include trade allocation
\end{enumerate}
\end{enumerate}
\end{process}

Hidden portions of large institutional orders are dark liquidity pools. Orders that are partially revealed are called iceberg or hidden-size orders.

\begin{remark}
\hlt{(Risks of Algorithmic Trading)}
\begin{enumerate}[label=\roman*.]
\setlength{\itemsep}{0pt}
\item Leaks might arise from competitor efforts to revere engineer them
\item Many algorithms lack the capacity to handle or respond to exceptional or rare events.
\end{enumerate}
\end{remark}

An auction is a competitive market process involving multiple buyers, multiple sellers, or both. Auctions are useful and cost-effective in pricing a security with an unknown value.

\begin{definition} {\color{white}space}
\begin{enumerate}[label=\roman*.]
\setlength{\itemsep}{0pt}
\item \hlt{English auction} is an ascending bid auction.
\item \hlt{Dutch auction} is a descending bid auction.
\item \hlt{First-Price Seal-Bid auctions} have all buyers simultaneously submit seal bids so that no bidders knows any of the other bids.
\begin{enumerate}[label=\arabic*.]
\setlength{\itemsep}{0pt}
\item Does not allow for price discovery until the auction concludes. \\
Winner submits highest bid and pays bid price.
\item Winner faces the Winners Curse problem if the auctioned item's value is not known with certainty.
\end{enumerate}
\item \hlt{Second-Price Seal-Bid (Vickrey) auctions}
\begin{enumerate}[label=\arabic*.]
\setlength{\itemsep}{0pt}
\item Identical to First-Price Sealed-Bid, except winner pays highest losing bid rather than his own bid.
\item Intended to encourage higher bids by reducing Winner's Curse.
\item Bidders bid more aggressively as a bid raises probability of winning without increasing expected cost, which is determined by someone else's bid.
\end{enumerate}
\item \hlt{Double auction} or \hlt{bilateral auction}
\end{enumerate}
\end{definition}

