\subsection{Market Fundmantals}

The basic function of a market is to bring buyers and sellers together.

\begin{process}
\hlt{(Four Components of a Trade)}
\begin{enumerate}[label=\roman*.]
\setlength{\itemsep}{0pt}
\item Acquisition of information and quotes
\begin{enumerate}[label=\arabic*.]
\setlength{\itemsep}{0pt}
\item Quality information and transparency are crucial to price discovery
\item Transparent markets quickly disseminate high-quality information
\item Opaque markets are those that lack transparency
\end{enumerate}
\item Routing of the trade order
\begin{enumerate}[label=\arabic*.]
\setlength{\itemsep}{0pt}
\item Selecting the brokers to handle the trades
\item Deciding which markets will execute the trades and transmitting the trades to the markets
\end{enumerate}
\item Execution. Buys are matched and executed against sells according to the rules of the market
\item Confirmation, clearance and settlement
\begin{enumerate}[label=\arabic*.]
\setlength{\itemsep}{0pt}
\item Clearance is the recording and comparison of the trade records
\item Settlement involves the actual delivery of the security and its payment
\item Might include trade allocation
\end{enumerate}
\end{enumerate}
\end{process}

\begin{remark}
\hlt{(Risks of Algorithmic Trading)}
\begin{enumerate}[label=\roman*.]
\setlength{\itemsep}{0pt}
\item Leaks might arise from competitor efforts to revere engineer them
\item Many algorithms lack the capacity to handle or respond to exceptional or rare events.
\end{enumerate}
\end{remark}

\begin{remark}
\hlt{(Cornering-the-Market)} Trader takes huge long futures position and tries to exercise control over supply of underlying commodity. As maturity of futures contract is approaching, position is not closed, number of outstanding contracts exceed commodities available. Holders of short positions desperately try to close positions, leading to rise in both futures and spot prices.\\
Abuse is dealt with by increasing margin requirements or imposing stricter position limits or prohibiting trades that increase speculator's open position or requiring market participants to close their positions.
\end{remark}

\subsubsection{Liquidity Access in Equity Markets}

\hlt{(Exchanges)} Account for $60\%$ to $70\%$ of all activity. The full order-book, arrivals/cancellations are all published, the liquidity information is transparent. Larger orders may impact the market. Liquidity at best price cannot be ignored, as the exchange will need to reroute to other exchanges where price is better while charging a fee (National Best Bid Offer, NBBO). All exchanges have almost exactly the same trading mechanism, and behave exactly the same way during the trading day except for opening and closing auctions.\\
Most exchanges use maker-taker fee model, but some exchanges (BYX, EDGA, NSX, BX) have an inverted fee model where rebate is provided for taking liquidity. IEX uses a speed bump to remove speed advantages of HFTs, providing a less 'toxic' liquidity pool.\\

\hlt{(Dark Pools/ATS)} Dark pools do not display any order information and use the NBBO as reference price. To avoid accessing protected venues, these pools trade only at the inside market (at or within the bid ask spread). Still possible to identify large blocks of liquidity by 'pinging' the pool at minimum lot size; counteracted by sending orders with minimum fill quantity tag, which allows block to be transparent from small pinging.\\
Most of ATS are run by major investment banks; some venues allow direct trading between investment firms. Note that many of the trading strategies used by firms tend to be highly correlated, hence liquidity is often on the same side; venues have to leverage sell side broker's liquidity to supplement their own.\\

\hlt{(Single Dealer's Platform/Systematic Internalisers)} Broker/Dealer and other institutional clients connect to Single Dealer Platform (SDP) directly. Not regulated ATS, hence can offer unique products. Brokers provide their own SDPs to expose their internal liquidity.\\

\hlt{(Auctions)} Exchanges begin and end day with a primary auction procedure that leverages special order types to accumulate supply and demand, then run algorithm that determines best price that would at best pair off the most volume. An opportunity for active and passive investors to exchange large amounts of liquidity.

\subsubsection{Trading Mechanisms}

Most common approach by modern electronic exchanges is the time/price priority, \hlt{continuous double auction} trading system. There are multiple buyers and multiple sellers participating at the same time.\\

\hlt{(Limit Order Book, LOB)} Stores all non-executed orders with associated instructions. Highly efficient, able to handle a high degree of concurrency to ensure the state is always correct. Has 2 copies of core data structure (for Buy, Sell orders). Supports 3 basic instruction types (insert, cancel, amend). There are 4 events on both Buy and Sell that may alter state of order book (limit order submission, limit order cancellation, limit order amendment, execution).\\

\hlt{(Matching Algorithm)} Responsible for interpreting various events to determine if any buy and sell orders can be matched. When multiple orders can be paired, price/time priority is used, where the order with most competitive prices are matched, and when prices are equal, the order that arrived prior is chosen.\\

Exchanges will publish order imbalance that exists among orders on opening and closing books during \hlt{Open Auctions}, with indicative price and volume. The following are published every second on market data feeds:
\begin{enumerate}[label=\roman*.]
\setlength{\itemsep}{0pt}
\item \hlt{Current Reference Price}: Price at which paired shares are maximised, imbalance is minimised, distance from bid-ask mid-point is minimised, in that order
\item \hlt{Near Indicative Clearing Price}: The crossing price at which orders in opening, closing book and continuous book would clear against each other
\item \hlt{Far Indicative Clearing Price}: The crossing price at which orders in opening, closing book would clear against each other
\item \hlt{Number of Paired Shares}: The number of on-open or on-close shares that is able to pair off at the current reference price
\item \hlt{Imbalance Shares}: The number of opening or closing shares that would remain unexecuted at the current reference price
\item \hlt{Imbalance Side}: The side of imbalance. B is buy-side imbalance; S is sell-side imbalance; N is no imbalance; O is no marketable on-open or on-close orders
\end{enumerate}

In general, the following rules apply to match supply and demand:
\begin{enumerate}[label=\roman*.]
\setlength{\itemsep}{0pt}
\item Crossing price must maximise volume transacted
\item If several prices result in similar volume transacted, the crossing price is the one the closest from last price
\item Cross price is identical for all orders executed
\item If two orders are submitted at same price, the order submitted first has priority
\item It is possible for an order to be partially executed if the other side quantity is not sufficient
\item 'At Market' orders are executed against each other at the determined crossing price, up to available matching quantity on both sides, but generally do not participate in price formation process
\item For Open Auction, unmatched 'At Market' orders are entered into continuous session of LOB as limit orders at the crossing price
\end{enumerate}

The open auction is a major price discovery mechanism, as it occurs after a period of market inactivity when market participants were unable to transact even if they have information. Market participants with better information are more likely to participate, with more aggressive orders to extract liquidity, hence the mechanism is quite volatile and more suited for short-term alpha investors.\\
During execution, smaller orders may avoid participation in open auction and period thereafter; larger orders may participate to extract significant liquidity from the market.\\

Diversity of order types is key component of continuous double auction electronic markets.

\begin{definition} {\color{white}space}
\begin{enumerate}[label=\roman*.]
\setlength{\itemsep}{0pt}
\item \hlt{Market Order}: trade carried out immediately at best price available in market.
\item \hlt{Limit Order}: only executed at this price or at one more favourable to the trader.
\item \hlt{Stop/Stop-Loss Order}: not visible in LOB. Only become active when certain price is reached or passed, then enter order book as ether limit or market order depending on user setup.
\item \hlt{Stop-Limit Order}: combination of stop order and limit order. Order becomes limit order as soon as a bid or ask is made at the price equal to or less favourable than stop price. If stop price and limit price is the same, then the order is \hlt{stop-and-limit} order.
\item \hlt{Trailing Stop Order}: function like stop orders, but stop price is set dynamic rather than static.
\item \hlt{Market-if-Touched (MIT)/Board Order}: executed at best available price after trade occurs at a specified price or more favourable. Ensure profit is taken if sufficiently favourable price movements occur.
\item \hlt{Market-Not-Held/Discretionary Order}: traded as market order, execution may be delayed at broker's discretion for better price.
\item \hlt{All-Or-None Order}: request full execution of order. Not executed until full quantity is available.
\item \hlt{Peg Order}: specify a price level at which order should be continuously and automatically repriced. Used for mid-point executions in non-displayed markets.
\item \hlt{Iceberg Order}: limit order with specific display quantity, designed to prevent information leakage.
\item \hlt{Hidden Order}: when available to trade, not directly available to other market participants in central LOB
\item \hlt{On-Open}: request execution at open price. Can be limit-on-open or market-on-open.
\item \hlt{On-Close}: request execution at close price. Can be limit-on-close or market-on-close.
\item \hlt{Imbalance Only}: provide liquidity intended to offset on-open/on-close order imbalances during opening/closing cross. These generally are limit orders.
\item \hlt{D-Quote}: Special order on NYSE, mainly used during close auction period.
\item \hlt{Funari}: Special order on Tokyo Stock Exchange, allows limit order placed in book during continuous session to automatically enter closing auction as market orders.
\end{enumerate}
\end{definition}

Instruction validity may take the following forms:
\begin{enumerate}[label=\roman*.]
\setlength{\itemsep}{0pt}
\item \hlt{Day Order}: valid for full duration of trading session
\item \hlt{Extended Day Orders}: allows for trading in extended hours
\item \hlt{Good-Till-Cancel Order}: in effect until executed or until end of trading in particular contract
\item \hlt{Immediate-or-Cancel Order}: will be immediately cancelled back to sender after reaching matching engine if it does not get immediate fill.
\item \hlt{Fill-or-Kill Order}: must be executed immediately on receipt or none at all.
\end{enumerate}

\subsubsection{Data Taxonomy}

Data availability, storage, management, and cleaning are the core of any research environment.\\

Reliable \hlt{reference data} is key foundation of a robust quantitative strategy development, with following data:
\begin{enumerate}[label=\roman*.]
\setlength{\itemsep}{0pt}
\item Trading Universe: evolving daily to incorporate new listings, de-listings etc. Knowing when a particular stock no longer trades is important to avoid survivor bias.
\item Symbology Mapping: ISIN, SEDOL, RIC, Bloomberg Tickers etc. Data is not static, symbols may change, complicating historical data merges. Mapping needs to persist as point-in-time data and allow for historical 'as-of-date' usage, require implementation of bi-temporal data structure. 
\item Ticker Changes: for reasons described in symbology mapping. To maintain historical table of ticker changes to seamlessly go up and down time series data.
\item Corporate Actions Calendars: contain stock and cash dividends (announcement, execution date), stock splits, reverse splits, rights offer, mergers and acquisitions, spin off, free float or shares outstanding adjustments, quotation suspensions etc.\\
For dividends, announcements may coincide with more volatility, jumps in price time series. Allow building of strategies that look to benefit from the added volatility.\\
For stock splits, reverse splits, rights offers, all historical data need to be adjusted backward to reflect the split (both volume and price).\\
For M\&A, spin-offs, to account for changes in valuation, hence used in Merger Arbitrage strategies.\\
Suspensions result in gaps in data, may impact backtesting.
\item Static Data: country, sector, primary exchange, currency, and quote factor. May be used to group instruments based on fundamental similarities (hence for pairs trading). Maintaining a table of quotation currency per instrument necessary to aggregate positions at portfolio level.
\item Exchange Specific Data: individual exchanges have variety of differences to be accounted for when designing trading strategies. First group of information concerns the hours and dates of operations:
\begin{enumerate}[label=\arabic*.]
\setlength{\itemsep}{0pt}
\item Holiday Calendar: Strategies trading simultaneously in several markets and leveraging correlation may not perform as well if one market is closed and another is open.
\item Exchange Sessions Hours: Different sessions (Pre-Market, Continuous Core, After-Hour etc.); auction times and respective cutoff times for oder submission; lunch break restricting intraday trading and auctions before/after lunch; settlement times for futures market. Daylight Saving Time (DST) adjustments; length of trading hours during course of the year; different trading hours by venue.
\item Disrupted Days: Exchange outages or trading disruptions, market data issues. To be recorded so they can be filtered out when building or testing strategies.
\end{enumerate}
Second group of information governing the mechanics of trading:
\begin{enumerate}[label=\arabic*.]
\setlength{\itemsep}{0pt}
\item Tick Size: Minimum eligible price increment; may vary by instrument and as a function of price.
\item Trade and Quote Lots: Minimum size increment for quotes or trades.
\item Limit-Up and Limit-Down Constraints: Maximum daily fluctuations of securities, and whether trading is paused or can only be traded at better prices than the threshold.
\item Short Sell Restrictions: Restrict shot sells not to trade at price worse than last price, or not to create a new quote that will be lower than the lowest prevailing quote. Impact ability to source liquidity.
\end{enumerate}
\item Market Data Condition Codes: vary per exchange and asset class, and each market event may be attributed to several codes at once. To build mapping table of condition codes and what they mean (i.e., auction trade, lit or dark trade, cancelled or corrected trade, regular trade, off-exchange trade reporting, block-size trade, trade originating from multi-leg order such as option spread trade etc.). To access liquidity for trading algorithm, trades published for reporting purposes must be excluded and not be used to update some of the aggregated daily data used in construction of trading strategies.
\item Special Day Calendars: days wth distinct liquidity characteristics to be accounted for in both execution strategies and in alpha generation process. These (non-exhaustive) irregular events may be:
\begin{enumerate}[label=\arabic*.]
\setlength{\itemsep}{0pt}
\item Half trading days preceding Christmas and following thanksgiving in US
\item Ramandan even in Turkey
\item Taiwan market opening on weekend to make up for lost trading days during holiday periods
\item Korean market changing trading hours on day of nationwide university entrance exam
\item Brazilian market opening late on day following the Carnival
\item Last trading days of months and quarters (investors rebalance portfolios)
\item Index rebalancing dates, where intraday volume distribution is significantly skewed toward EODs
\item Options and futures expiry dates (quarterly/monthly expiry, Triple Witching in US, Special Quotations in Japan) where excess trading volume and different intraday patterns result from hedging activity and portfolio adjustments.
\end{enumerate}
Model normal days first. Special days are modelled either independently, or using normal days as baseline.
\item Futures-Specific Reference Data: to know which contract was live at any point of time by using expiry calendar, and the most liquid contract. Equity index futures are most liquid for first contract available (front month), energy futures such as oil are more liquid for second contract. Hence to know which contract carry the most significant price formation characteristics, and what is true liquidity available.\\
Note there is no real standardised expiry frequency that applies across markets. When computing rolling-window metrics, to account for potential roll dates (due to investors rolling forward positions) that may have happened during the time span. May blend volume time series prior to roll date and after roll date.\\
Futures market also have different market phases during the day with significantly different liquidity characteristics. Various market data metrics (volume profile, average spread, average bid-ask sizes etc) should be computed separately for each market phase by maintaining a table of start and end times of each session for each contract.
\item Options-Specific Reference Data (Options Chain): expiry date and strike price combination (option chain). Map of equity tickers to option tickers with strike and expiry dates allow for design for more complex investment and hedging strategies (i.e., distance to strike, change in open interest of puts and calls).
\item Market-Moving News Releases: macroeconomic announcements. To maintain calendar of dates and times of their occurrences to assess their impact on strategies. Central bank announcements or meeting minutes releases about major economies (FED/FOMC, ECB, BOE, BOJ, SNB), Non-Farm Payrolls, Purchasing Managers' Index, Manufacturing Index, Crude Oil Inventories etc.\\
Stock-specific releases such as earnings calendars, specialised sector events (for healthcare, biotech etc).
\item Related Tickers: tickers that are related to each other as they fundamentally represent the same underlying asset. Allows efficient opportunity exploitation. Primary tickers to composite tickers mapping (for markets with fragmented liquidity), dual listed/fungible securities in US and Canada, ADR or GDR, local and foreign boards in Thailand etc.
\item Composite Assets: ETFs, Indexes, Mutual Funds etc. May be used to achieve desired exposures, or as cheap hedging instruments, and provide arbitrage opportunities when they deviate from NAV. To maintain information such as time series of their constituents and value of any cash component, divisor used to translate NAV into quoted price, constituent weights.
\item Latency tables: for higher frequency trading strategies. Contains distribution of latency between different data centres for more efficient order routing, and reordering data that are recorded in different locations.
\end{enumerate}

There are different details in the \hlt{market data} feed, as follows:
\begin{enumerate}[label=\roman*.]
\setlength{\itemsep}{0pt}
\item Level I Data (Trade and BBO Quotes): trades and top of book quotes. Enough to reconstruct Best Bid and Offer (BBO). Also contains information in form of trade status (cancelled, reported late etc), trade and quote qualifiers (odd lot, normal trade, auction trade, Intermarket Sweep, average price reporting, on which exchange etc). May be used to analyse sequence of events and decide if a given print should be used to update the last price and total volume traded at a point in time.
\item Level II Data (Market Depth): addition of quote depth data, displays all lit limit order book updates (price changes, addition or removal of shares quoted) at any level in the book, for all of the lit venues in fragmented markets.
\item Level III Data (Full Order View): message data. Each order arriving is attributed a unique ID for tracking over time, and is precisely identified when it is executed, cancelled, or amended. Possible to build a full (with national depth) book at any moment intraday. Example from US market:
\begin{enumerate}[label=\arabic*.]
\setlength{\itemsep}{0pt}
\item Timestamp: number of milliseconds after midnight
\item Ticker: equity symbol (up to 8 char)
\item Order: Unique order ID
\item T: message type. 'B' is add buy order; 'S' is add sell order'; 'E' is execute outstanding order in part; 'C' is cancel outstanding order in part; 'F' is execute outstanding order in full; 'D' is delete outstanding order in full; 'X' is bulk volume for cross event; 'T' is execute non-displayed order
\item Shares: order quantity for 'B', 'S', 'E', 'X', 'C', 'T' messages. Zero for 'F', 'D' messages
\item Price: order price for 'B', 'S', 'X', 'T' messages. Zero for cancellation and executions. Last 4 digits are decimal, padded to right with zeroes. Divide by $1000$ to convert to currency value.
\item MPID: Market Participant ID associated with transaction (4 char)
\item MCID: Market Centre Code for originating exchange (1 char)
\end{enumerate}
A few special types of orders worth mentioning are:
\begin{enumerate}[label=\arabic*.]
\setlength{\itemsep}{0pt}
\item Order subject to price sliding: execution price may be one cent worse than display price at NASDAQ; ranked at locking price as hidden order, displayed at the price one minimum price variation inferior to locking price. New order ID will be used if order is replaced as a display order.
\item Pegged order: based on NBBO, not routable, new timestamp given upon repricing; display rule vary over exchanges
\item Mid-point peg order: non-displayed, may result in half-penny execution
\item Reserve order: displayed size is ranked as displayed limit order; reserve size is behind non-displayed orders and pegged orders in priority.\\
Minimum display quantity is 100, amount replenished from reserve size when it falls below 100 shares; New timestamp created, displayed size re-ranked upon replenishment.
\item Discretionary order: displayed at one price while passively trading a more aggressive discretionary price. Order becomes active when shares are available within discretionary price range. Order ranked last in priority. Execution price may be worse than display price.
\item Intermarket sweep order: can be executed without need for checking prevailing NBBO. 
\end{enumerate}
Using these data, we may model: the pattern of inter-arrival times of various events; arrival and cancellation rates as a function of distance from nearest touch price; arrival and cancellation rates as a function of other information, such as in the queue on either side of the book, order book imbalance etc.\\
Once modelled, we may analyse: the impact of market order on limit order book; chances for limit order to move up the queue from given entry position; probability of earning the spread; expected direction of price movement over a short horizon.
\end{enumerate}

Most quantitative strategies rely on \hlt{binned data}, such as the following daily statistics:
\begin{enumerate}[label=\roman*.]
\setlength{\itemsep}{0pt}
\item Open, High, Low, Close (OHLC) and Previous Close Price: indication on trading activity and intraday volatility. Distance traveled between lowest and highest points is indication of market sentiment. Previous close has to be adjusted for corporate actions and dividends.
\item Last Trade before Close (Price/Size/Time): how much the close price may have jumped in final moments of trading; how stable it is as a reference value for next day.
\item Volume: trading activity indicator, especially when level jumps from long term average. Collect volume breakdown between lit and dark venues for execution strategies.
\item Auctions Volume and Price: price discovery event when significant volume prints occur.
\item VWAP: indication of trading activity on the day. Easier to algorithmically execute large orders with VWAP than a single print.
\item Short Interest/Days-to-Cover/Utilisation: good proxy for investor position. Short pressure an indication of upcoming short term moves: large short interest is bearish view from institutional investors. Utilisation level of available securities to borrow gives indication of how much room is left for future shorting. Days-to-Cover to assess magnitude of potential short squeeze (if sellers unwind position, fraction of available daily liquidity needed); larger value indicates larger potential of sudden upswing on heavily shorted securities.
\item Futures Data: insight into activity or large investors through open interest data. Offer arbitrage opportunities if their basis exhibits mis-pricing compared to one's dividend estimates.
\item Index-Level Data: source of relative measures for instrument specific features (index OHLC, volatility). Normalised features identify individual instruments deviating from their benchmarks.
\item Options Data: information on position of traders through open interest and Greeks.
\item Asset Class Specific: yield/benchmark rates (repo, 2y, 10y, 30y), CDS spreads, US Dollar Index
\end{enumerate}
We also consider data for granular intraday microstructure activity:
\begin{enumerate}[label=\roman*.]
\setlength{\itemsep}{0pt}
\item Number and Frequency of Trades: proxy for activity level, and how continuous it is. Low number of trades mean harder execution, and may be more volatile
\item Number and Frequency of Quote Updates: similar proxy for activity level
\item Top of Book Size; proxy for liquidity of instruments (larger top of book size makes it possible to trade larger order quasi immediately)
\item Depth of Book (price and size): similar proxy for liquidity
\item Spread Size (average, median, time weighted average): proxy for cost of trading. Parametrised distribution used to identify opportunities if they are cheap or expensive
\item Trade size (average, median): to identify intraday liquidity opportunities when examining volume available in the order book.
\item Ticking time (average, median): representation of how often one should expect changes in the order book first level. For execution algorithms for which the frequency of updates (adding/cancelling child orders, re-evaluating decisions etc.) should commensurate with characteristics of the traded instrument.
\end{enumerate}
Daily distribution of these intraday variables can be used as start of day estimates and updated intraday with online Bayesian updates.\\
Last group of daily data is derived from previous two groups but stored pre-computed to save time during research phase, or to be used as normalising values:
\begin{enumerate}[label=\roman*.]
\setlength{\itemsep}{0pt}
\item X-day Average Daily Volume (ADV) / Average Auction Volume
\item X-day Volatility (close-to-close, open-to-close etc)
\item Beta with respect to index or sector (plain beta, or asymmetric up-days/down-days beta)
\item Correlation matrix
\end{enumerate}

When binning data, this may be grouped into bins ranging from a few seconds to 30 minutes. Minute bar data are used for volume and spread profiles to prevent introducing excess noise due to market friction.\\

Some \hlt{fundamental data} and other data sets which may be considered are as follows:
\begin{enumerate}[label=\roman*.]
\setlength{\itemsep}{0pt}
\item Key Ratios: Earnings Per Share (EPS), Price-to-Earning (P/E), Price-to-Book (P/B), etc. 
\item Analyst Recommendations: aggregated values given consensus valuation
\item Earnings data: estimations by research analysts provide quarterly earning estimates which can be used as indication of performance of stock before actual value is published
\item Holders: sudden changes in ownership indicate changes in sentiment by sophisticated investor
\item Insiders Purchase/Sale: indicator of future stock price moves from group of people who have access to best possible information about the company
\item Credit Ratings: credit downgrades resulting in higher funding costs in the future have negative impact on equity prices
\end{enumerate}

\subsubsection{Market Microstructure Primer}


